---
title: "Class 10"
author: "Diana Lopez"
format: gfm
editor: visual
---

# Importing candy data

```{r}
candy_url <- url("https://raw.githubusercontent.com/fivethirtyeight/data/master/candy-power-ranking/candy-data.csv")
candy <- read.csv(candy_url, header = TRUE, row.names=1)

head(candy)
```

Q1. How many different candies are in this dataset?

85

Q2. How many fruity candy types are in this data set?

```{r}
sum(candy[,2])
```

# What is your favorite candy?

```{r}
candy["Root Beer Barrels", ]$winpercent

candy["Kit Kat", ]$winpercent

candy["Tootsie Roll Snack Bars", ]$winpercent
```

```{r}
library("skimr")
skim(candy)
```

-   **Q6**. Is there any variable/column that looks to be on a different scale to the majority of the other columns in the dataset?

The winpercent is a different scale because the mean is 50 vs \~0-1

-   **Q7**. What do you think a zero and one represent for the `candy$chocolate` column?

The zero vs one represents the identity of a candy being or containing chocolate.

-   **Q8**. Plot a histogram of `winpercent` values

```{r}
hist(candy[,"winpercent"])
```

-   **Q9**. Is the distribution of `winpercent` values symmetrical?

    No

<!-- -->

-   **Q10**. Is the center of the distribution above or below 50%?

    Below 50%

<!-- -->

-   **Q11**. On average is chocolate candy higher or lower ranked than fruit candy?

```{r}
chocolate.inds <- as.logical(candy$chocolate)
chocolate.win <- candy[chocolate.inds,]$winpercent
mean(chocolate.win)
fruity.inds <- as.logical(candy$fruit)
fruity.win <- candy[fruity.inds,]$winpercent
mean(fruity.win)
```

-   **Q12**. Is this difference statistically significant?

```{r}
t.test(chocolate.win,fruity.win)
```

-   **Q13**. What are the five least liked candy types in this set?

```{r}
ordered_by_winpercent <- candy[order(candy[,"winpercent"]),]
head(ordered_by_winpercent, n=5)
library("dplyr")
candy %>% arrange(winpercent) %>% head(5)
```

-   **Q14**. What are the top 5 all time favorite candy types out of this set?

```{r}
ordered_by_winpercent <- candy[order(candy[,"winpercent"],decreasing = TRUE),]
head(ordered_by_winpercent, n=5)
library("dplyr")
candy %>% arrange(winpercent) %>% head(5)
```

-   **Q15**. Make a first barplot of candy ranking based on `winpercent` values. **Q16**. This is quite ugly, use the `reorder()` function to get the bars sorted by `winpercent`?

```{r}
library("ggplot2")
ggplot(ordered_by_winpercent)+
  aes(winpercent,rownames(candy))+
  geom_col()
```

```{r}
library("ggplot2")
ggplot(candy)+
  aes(winpercent,reorder(rownames(candy),winpercent))+
  geom_col()
```

-   **Q17**. What is the worst ranked chocolate candy?

```{r}
my_cols=rep("black",nrow(candy))
my_cols[as.logical(candy$chocolate)] <- "chocolate"
my_cols[as.logical(candy$bar)] <- "brown"
my_cols[as.logical(candy$fruity)] <- "pink"
```

```{r}
ggplot(candy)+aes(winpercent, reorder(rownames(candy),winpercent))+
  geom_col(fill=my_cols)+theme(text=element_text(size=6))

ggsave("mybarplot_withcol.png")
```

-   Q18. What is the best ranked fruity candy

    Star bursts

```{r}
library(ggrepel)

ggplot(candy)+
  aes(winpercent, pricepercent, label=rownames(candy))+
  geom_point(col=my_cols)+
  geom_text_repel(col=my_cols, size=3.3, max.overlaps=5)
```

-   **Q19**. Which candy type is the highest ranked in terms of `winpercent` for the least money - i.e. offers the most bang for your buck?

    Reese's miniatures

<!-- -->

-   **Q20**. What are the top 5 most expensive candy types in the dataset and of these which is the least popular?

```{r}
ordered_by_pricepercent <- candy[order(candy[,"pricepercent"],decreasing = TRUE),]
head(ordered_by_pricepercent, n=5)
```

# Exploring the correlation structure

```{r}
library(corrplot)
cij <- cor(candy)
corrplot(cij)
```

-   **Q22**. Examining this plot what two variables are anti-correlated (i.e. have minus values)?

    Fruity with chocolate and pluribus with bar

<!-- -->

-   **Q23**. Similarly, what two variables are most positively correlated?

    Chocolate with winpercent, and chocolate with Bar

# Principal Component Analysis

```{r}
candy_pca <- prcomp(candy,scale=TRUE)
summary(candy_pca)
```

```{r}
plot(candy_pca$x[,1],candy_pca$x[,2],col=my_cols,pch=16)
```

```{r}
my_data <- cbind(candy,candy_pca$x[,1:3])
```

```{r}
plot_1 <- ggplot(my_data)+
              aes(x=PC1, y=PC2,
                  size=winpercent/100,
                  text=rownames(my_data),
                  label=rownames(my_data))+
              geom_point(col=my_cols)+
              geom_text_repel(size=3.3,col=my_cols,max.overlaps = 7)+
  labs(title="Halloween Candy PCA Space",
       subtitle="Colored by type: chocoolate bar (dark brown), chocolate other (light brown), fruity (pink)",
       caption="data from 538 points")
plot_1
```

-   **Q24**. What original variables are picked up strongly by PC1 in the positive direction? Do these make sense to you?

```{r}
barplot(candy_pca$rotation[,1], las=2,ylab="PCA1 contribution")
```

```{r}
barplot(candy_pca$rotation[,2], las=2,ylab="PCA1 contribution")
```

```{r}

x <- summary(candy_pca)
plot(x$importance[2,],type="b")
```
